Behavior Driven Development in Haskell

### 2012-05-23 paper.tex
writing observation.

@@

Why do we dislike to write test programs?
It is probably because writing test programs is not fun.
To encourage Haskellers to write more test programs
to detect value errors,
we need a scheme to make writing test programs fun.

@@

In our method, the text for design takes role of documentation and
is used as test cases. 
This method would motivate Haskellers
bacause it has a kill-three-birds-with-one-stone aspect.

@@

\begin{description}
\item[unit test] TBD
\item[example] TBD
\item[property] TBD
\item[behavior] TBD
\item[documentation] Text written in Haskell comments and parsed by Haddock
\end{description}

@@

Haddock example and doctest.

Should talk about equation property. (e.g. container)
Quickcheck.

pure vs IO.

@@

We would like to thank
Nick Smallbone for having improved QuickCheck so that boolean property is tested
just once.
We are grateful to David Waern and Simon Marlow
for introducing the property syntax to haddock.

>>>>

TBD: we need evidence to support above.

@@

Why do we dislike to write test programs?
It is probably because writing test programs is not fun.
To encourage Haskellers to write more test programs
to detect value errors,
we need a scheme to make writing test programs more interesting.

@@

In our method, design specification takes role of documentation for users and
is used as test cases by its developer. 
This method would motivate Haskellers
bacause it has a kill-three-birds-with-one-stone aspect.

@@

The definitions of words used in this paper is as follows:

\begin{description}
\item[unit test] tests for individual programs or libraries.
\item[example] Typical usage of functions.
\item[property] Invariants which functions have.
\item[behavior] How do functions behave for users. Behavior can be shown by using both example and property.
\item[design] TBD
\item[documentation] Explanation of functions.
\end{description}

@@

We insist that examples and properties, which are useful for users, should be
documented.
So, let us first observe Haddock, the documentation system of Haskell. Haddock defines annotation syntax in Haskell's comment.

\subsection{Example annotation}

We can specify examples with the annotation syntax {\tt >>>}.
The following is an example of how to use the {\tt length} function
for lists.

\begin{quote}
\begin{verbatim}
> length []
0
\end{verbatim}
\end{quote}

\noindent We can write this by copy-and-pasting from an interpreter.
It's worth explaining that we can specify operation sequences:

\begin{quote}
\begin{verbatim}
> writeFile "tmpfile" "Hello"
> readFile "tmpfile"
"Hello"
\end{verbatim}
\end{quote}

\noindent But example annotations are useful for users but 
it is not used as test case from the developer point of view (before we implemented {\tt doctest}).

\subsection{Property annotation}

The {\tt container} packages has equation properties:

\begin{quote}
\begin{verbatim}
>>> Data.Map.null (empty) == True
\end{verbatim}
\end{quote}

It is written in code blocks ({\tt >}).
This cannot be used as test case.

Equation property is tested 100 times.

Quickcheck property should be able to be specified.

To our experience, it is not difficult to find properties for pure functions.
We believe that Haskellers should show properties of pure functions
in documentation.

\subsection{Internal documentation}

not useful for users

@@

We would like to thank
Nick Smallbone for having improved QuickCheck so that boolean property is tested
just once.
We are grateful to David Waern and Simon Marlow
for introducing the property syntax to Haddock.
